---
title: "DATA 606 Chapter 5 Homework"
author: "Yohannes Deboch"
date: "March 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
5.6 Working backwards, Part II.
A 90% confidence interval for a population mean is (65,77). The population distribution is approximately normal and the population standard deviation is unknown. This confidence interval is based on a simple random sample of 25 observations. Calculate the sample mean, the margin of error, and the sample standard deviation.

Answer:
```{r}
n <- 25
margin_error <- (77-65)/2
margin_error
t_value <- qt(0.95,n-1)
mean <- (77+65)/2
mean
sd <- (margin_error/t_value)*5
sd
```
Mean 71, margin of error 6 and sample standard deviation  17.53481.

5.14 SAT scores. SAT scores of students at an Ivy League college are distributed with a
standard deviation of 250 points. Two statistics students, Raina and Luke, want to estimate the
average SAT score of students at this college as part of a class project. They want their margin of
error to be no more than 25 points.

(a) Raina wants to use a 90% confidence interval. How large a sample should she collect?

Answer:
```{r}
z <- qnorm(0.95)
n <- (z*250/25)^2
n
```
The sample size should be  271.

(b) Luke wants to use a 99% confidence interval. Without calculating the actual sample size,
determine whether his sample should be larger or smaller than Raina’s, and explain your
reasoning.

Answer:
Luke's sample size will be larger than Raina's since increasing confidence interval will increase the value of Z in the numerator of sample size calculation formula which will in turn increase the sample size.

(c) Calculate the minimum required sample size for Luke.


Answer:
```{r}
z <- qnorm(0.01/2)
n <- (z*250/25)^2
ceiling(n)
```
The Luke's minimum sample size should be  664


5.20 High School and Beyond, Part I. The National Center of Education Statistics conducted
a survey of high school seniors, collecting test data on reading, writing, and several other subjects.
Here we examine a simple random sample of 200 students from this survey. Side-by-side box plots
of reading and writing scores as well as a histogram of the differences in scores are shown below.

(a) Is there a clear difference in the average reading and writing scores?


Answer:
There is not clear difference in the average reading and writing scores since the distribution of the difference scores is approximately symmetrical.


(b) Are the reading and writing scores of each student independent of each other?

Answer: The reading and writing scores of each students are independent of each other.

(c) Create hypotheses appropriate for the following research question: is there an evident difference
in the average scores of students in the reading and writing exam?

Answer: 
Null Hypothesis: There is no difference between mean writing and reading score of the students.
Alternative Hypothesis: There is  difference between mean writing and reading score of the students.

(d) Check the conditions required to complete this test.

Answer: The following conditions are true
1. Observations are independent
2. Difference of scores of Observations are approximately normal and the sample size is large.

(e) The average observed difference in scores is ¯ x read−write = −0.545, and the standard deviation
of the differences is 8.887 points. Do these data provide convincing evidence of a difference
between the average scores on the two exams?

Answer:
```{r}
se <- 8.887/sqrt(200)
t_statistics <-  −0.545/se
t_statistics
print("two tailed p-value")
pt(t_statistics, 200-1)*2
```
Since the p value is > 0.05 at 5% level of significance we accept the null hypothesis.We can say that 
these data don't provide convincing evidence of a difference
between the average scores on the two exams.

(f) What type of error might we have made? Explain what the error means in the context of the
application.

Answer: We might have made type 2 error. The mean difference of the reading and writing score is different than 0.

(g) Based on the results of this hypothesis test, would you expect a confidence interval for the
average difference between the reading and writing scores to include 0? Explain your reasoning.


Answer:I would expect a confidence interval for the average difference between the reading and writing scores to include 0.Because the sample estimate is close to the null estimate.
Confidence interval
```{r}
−0.545+t_statistics
−0.545-t_statistics
```

5.32 Fuel efficiency of manual and automatic cars, Part I. Each year the US Environs-
mental Protection Agency (EPA) releases fuel economy data on cars manufactured in that year.
Below are summary statistics on fuel efficiency (in miles/gallon) from random samples of cars with
manual and automatic transmissions manufactured in 2012. Do these data provide strong evidence
of a difference between the average fuel efficiency of cars with manual and automatic transmissions
in terms of their average city mileage? Assume that conditions for inference are satisfied. 

Answer: 
Null Hypothesis: The data provides strong evidence that there is no  difference between the average fuel efficiency of cars with manual and automatic transmissions
in terms of their average city mileage
Alternative Hypothesis:

The data provides strong evidence that there is  difference between the average fuel efficiency of cars with manual and automatic transmissions
in terms of their average city mileage
```{r}
t_statistics <- (19.85-16.12)/sqrt((4.51^2+3.58^2)/26)
t_statistics
# two_tailed p value
pt(t_statistics,26-1, lower.tail = FALSE)*2
```
Since the p value is 0.002883615 < 0.05 at 5% we reject the null hypothesis and conclude that the  data provides strong evidence that there is difference between the average fuel efficiency of cars with manual and automatic transmissions
in terms of their average city mileage

5.48 Work hours and education. The General Social Survey collects data on demographics,
education, and work, among many other characteristics of US residents. 47 Using ANOVA, we can
consider educational attainment levels for all 1,172 respondents at once. Below are the distributions
of hours worked by educational attainment and relevant summary statistics that will be helpful in
carrying out this analysis.


(a) Write hypotheses for evaluating whether the average number of hours worked varies across the
five groups.

Answer: Null Hypothesis: Average weekly exercise hour is same for all the groups
Alternative Hypothesis:
Average weekly exercise hour is not same for all the groups
(b) Check conditions and describe any assumptions you must make to proceed with the test.
Answer: 
The observations are independent.

Observations within each groups are approximately normal we can see from the box plot except HS and Bachelor's seem to exhibit newness. As each group size is large it won't be a big problem.

Variance of the 5 groups are approximately similar.


(c) Below is part of the output associated with this test. Fill in the empty cells.
Df Sum Sq Mean Sq F value Pr(>F)
degree XXXXX XXXXX 501.54 XXXXX 0.0682
Residuals XXXXX 267,382 XXXXX
Total XXXXX XXXXX

Answer: 
```{r}
n <- 1172
group <- 5
MSG <- 501.54
SSE <- 267382
(df_group <- group -1)
(df_R <- n - group)
(df_total <- n-1)
(SSG <- MSG * df_group)
(SST <- SSG + SSE)
(MSE <- SSE/df_R)
(F <- MSG / MSE)

data.frame(X=c("degree","Residuals","Total"),
           DF = c(df_group, df_R,df_total),
           "Sum Sq" =c(SSG,267382,SST),
           "Mean SQ" = c(501.54,MSE," "),"F value" = c(F,NA,NA),"Pr(>F) "= c(0.0682,NA,NA))
```


(d) What is the conclusion of the test?

Answer: Since the P value 0.0682 > 0.05 at 5% level of significance we conclude that there is no significance difference among the groups for average weekly physical exercise.
